---
title: "Get Baysian score"
author: "Hoor Al-Hasani"
date: "2026-03-23"
params:
  epitox: "~/Documents/Projects/MAGEA3/EpiTox_results/Cutoff_4/Table/HPA_genes_nTPM.xlsx"   
  xscan: "~/Documents/Projects/MAGEA3/EpiTox_results//Cutoff_4/Table/Xscan_MAGEA3.xlsx"
  iedb_files: "~/Documents/Projects/MAGEA3/EpiTox_results/Cutoff_4/PrediTopes/DB_annotation/"
  target_allele: "HLA-A*02:01"
  target: "KVAELVHFL"
  affinity: 14.39613
  DIR: "./data/Bayesian/MAGEA3" # with prefix
  SaveResults: false
  
output: html_document
---

# Declare libraries

```{r setup, include=FALSE, echo=FALSE, comment="#"}

biocopy_colors = c("#A2C510", "#99CFE9", "#FBB800", "#939597", "#C61E19", "#438D99", "#958BB2", "#6B7B88",
                   "#338232", "#F08000", "#3373A1", "#64686A", "#D14B47", "#98D0BC", "#4F3D7F", "#2C4255")


knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lessR)
library(tidyr)
library(patchwork)

#library(ggplot2)
#library(ggpubr)
#library(Peptides)
#library(tidyverse)
#library(corrplot)
#library(ggrepel)
```

### Reading input

the input to the model is all the annotations that EpiTox extracted from IEDB. For this target (_MAGEA3_), both HLA-ligand atlas and PEPREP contribution was subset of the IEDB whole annotation. See notes bellow to understand some of the decisions that were made.

The input is:

- the final table from V2.5's `annotopes`: _HPA_genes_nTPM.xlsx_

- the experimental evidence file from the previous step

- _IEDB_ annotated peptides


## Step 1: Estimate _Experimental Evidence_

The experimental evidence is a Baysian-based model to evaluate and summarize the context evidence that is available for each peptide.
the concept is to determine the peptide with its protein ensuring context aware evaluation.


```{r iedb, include=T, echo=F, comment=NA}

# Bayesian_LR_sensitivity.R: to access LR thresholds
# Bayesian_workflow.R: the main orchestrator -> will be tranformed to this section directly
# utils_exp.R: will test if it is needed at all
# Bayesian_peptide_assessment.R:
# Bayesian_preprocess_experimental_data.R:

source("../MAGEA3_analysis/utils_exp.R")
source("../MAGEA3_analysis/Bayesian_peptide_assessment.R")
source("../MAGEA3_analysis/Bayesian_preprocess_experimental_data.R")

files = list.files(path = params$iedb_files,
                   pattern = "*sequence.csv", full.names = T)

length(files) %>% paste(., "files were found!") %>% print()
basename(files) %>% print()

experimental_df = lapply(files, function(f){
  if (file.size(f) == 0) return(NULL)
  first_line <- readLines(f, n = 1, warn = FALSE)
  if (length(first_line) == 0 || all(trimws(first_line) == "")) return(NULL)

  x = read.csv(f) %>%
    select(parent_source_antigen_name, curated_source_antigen.accession, linear_sequence, reference_id,
           qualitative_measure, mhc_class, mhc_allele_name,
           assay_names, disease_names) %>%
    dplyr::rename(
      peptide = linear_sequence,
      study_ref = reference_id,
      disease = disease_names,
      Experimental_method = assay_names,
      Binding = qualitative_measure,
      hla_class = mhc_class,
      hla_allele = mhc_allele_name)
  return(x)

}) %>%
  do.call("rbind",.) %>%
  mutate(uniprot = gsub("\\.\\d+", "", curated_source_antigen.accession),
         disease = gsub("\\[\\'(.+)\\'\\]", "\\1", disease),
         curated_source_antigen.accession = NULL,
         id = paste0(uniprot, "_", peptide),
         uniprot = NULL, peptide = NULL,
         is_normal = case_when(
           disease == "healthy" ~ TRUE,
           TRUE ~ FALSE
         )) %>%
  relocate(id)
```

```{r epitox, include=FALSE, echo=FALSE, comment="#"}
# loading EpiTox's peptide list
# --------------------------------
epitox = openxlsx::read.xlsx(params$epitox) %>%
  select(ensembl_gene_id, uniprot, Gene.Names, peptide, blosum_similarity, mismatch,
         Wildtype, Peptide_HLA_Atlas, affinity, presentation_score) %>%
  distinct(.keep_all = T) %>%
  mutate(id = paste0(uniprot, "_", peptide)) %>%
  dplyr::rename()%>%
  distinct(peptide, uniprot, .keep_all = T)

```

### Core analysis

process experimental data, followed by bayesian assessment then saving the results

```{r preprocess, include=FALSE, echo=FALSE, comment="#"}
preprocessed <- preprocess_for_bayesian(epitox, experimental_df,
                                        target_allele = params$target_allele,
                                        target_disease = NULL,
                                        target_class = "I")

# Assess
results <- bayesian_peptide_assessment(
  preprocessed,
  target_allele = params$target_allele
)

if (params$SaveResults) generate_assessment_report(results, output_prefix = params$DIR)

```

#### Generate plots
```{r visualize_exp_pie, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=15}

cols <- c("High" = "#C61E19", "Medium" = "#FDE399",
          "Low" = "#8ECAE6", "Very Low" = "#A2C510")

df_pie1 <- as.data.frame(table(results$confidence_level))
colnames(df_pie1) <- c("confidence_level", "count")
p1 <- ggplot(df_pie1, aes(x = "", y = count, fill = confidence_level)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = cols) +
  labs(title = paste0("Total peptides: ", nrow(results)), fill = "Confidence level") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

df_pie2 <- as.data.frame(table(results$interpretation))
colnames(df_pie2) <- c("interpretation", "count")
p2 <- ggplot(df_pie2, aes(x = "", y = count, fill = interpretation)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = biocopy_colors) +
  labs(title = paste0("Total peptides: ", nrow(results)), fill = "Interpretation") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p1 + p2
```


```{r visualize_I, echo=FALSE, comment=NA, fig.keep='all', results = 'hide', warning=FALSE, , message=FALSE}
g = ggplot(results, aes(affinity/10^5, posterior_prob)) +
  geom_point(size = 1, aes(col = evidence)) +
  theme_light()

#print (g)

g =ggplot(results, aes(posterior_prob, affinity*10^-5)) +
  geom_point(size = 1, aes(col = confidence_level)) +
  theme_light() +
  facet_wrap(~confidence_level)

#print(g)
```

## Step 2: TPT

### Target

```{r TPT, include=T, echo=F, comment=NA}
# create TPT -> target pattern/profile template
source("../MAGEA3_analysis/xscan_target_utils.R") # for target functions
source("../MAGEA3_analysis/xscan_OT_utils.R") # for OT functions


#target = data.frame(peptide = params$target, params$affinity)
substitution_matrix <- openxlsx::read.xlsx(params$xscan) %>%
  select(peptide, affinity) %>%
  convert_to_matrix(df = ., wt_peptide = as.character(params$target))

# Run analysis
anchor_results <- analyze_anchor_positions(substitution_matrix, as.numeric(params$affinity), params$target)
anchor_results <- analyze_position_compensation(anchor_results)

```

```{r visualize_II, echo=FALSE, comment=NA, fig.keep='all', results = 'hide', warning=FALSE, , message=FALSE, fig.height=5, fig.width=15}
L = nchar(params$target)
p1 = ggplot(anchor_results, aes(x = Position, y = Fold_Change_Mean)) +
  geom_col(aes(fill = Category), color = "#2C4255") +  # Changed from Is_Anchor to Category
  geom_hline(yintercept = 1, linetype = "dashed", color = "#A2C510") +
  scale_fill_manual(values = c("Classical Anchor" = "#A2C510",
                               "Permissive" = "#99CFE9",
                               "Flexible Anchor" = "#4F3D7F",
                               "Neutral" = "#F0F0F0"),
                    name = "Position Type") +
  labs(title = "Mean Fold Change in KD by Position",
       x = "Peptide Position",
       y = "Mean Fold Change\n(Mutant KD / WT KD)") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        legend.position = "none") +
  scale_x_continuous(breaks = 1:L)



p2 = ggplot(anchor_results, aes(x = Position, y = Percent_Better_Than_WT)) +
    geom_col(aes(fill = Category), color = "#2C4255") +
    geom_hline(yintercept = 50, linetype = "dashed", color = "#2C4255") +
    scale_fill_manual(values = c("Classical Anchor" = "#A2C510",
                               "Permissive" = "#99CFE9",
                               "Flexible Anchor" = "#4F3D7F",
                               "Neutral" = "#F0F0F0"),
                    name = "Position Type") +
    labs(title = "Position Permissiveness",
       subtitle = "% of substitutions that maintain/improve binding",
       x = "Peptide Position",
       y = "% Better or Equal to WT") +
    theme_light() +
    theme(text = element_text(size = 10),
                          legend.position = "bottom") +
    scale_x_continuous(breaks = 1:L) +
    coord_flip()

p1 +p2

```

### Off-target classification

```{r TPT_OT, include=T, echo=F, comment=NA}
target_classifications = anchor_results %>%
  mutate(WT_AA = strsplit(params$target, "")[[1]]) %>%
  select(Position, WT_AA, Category)

offtargets_data = epitox %>%
  select(uniprot, peptide, Wildtype, blosum_similarity, mismatch,
         affinity, presentation_score) %>%
  distinct(.keep_all = T)

# Perform complete analysis
analysis_results <- analyze_offtargets(params$target, target_classifications, offtargets_data)

# Create comprehensive summary table
summary_table <- create_summary_table(analysis_results) %>%
  mutate(uniprot = gsub(".+\\_", "", id),
         peptide = gsub("\\_.+$", "", id),
         id = NULL) %>%
  relocate(uniprot, peptide) %>%
  arrange(Total_Mismatches)

categorized <- summary_table %>%
  mutate(
    Category = case_when(
      Anchor_Mismatches > 0 & Exposed_Mismatches > 0 ~ "Both Anchor & Backbone",
      Anchor_Mismatches > 0 & Exposed_Mismatches == 0 ~ "Anchor Only",
      Anchor_Mismatches == 0 & Exposed_Mismatches > 0 ~ "Backbone Only",
      TRUE ~ "Perfect Match"
    )
  )

categorized %>% head %>% print()
```

```{r visualize_mutliFeatures_pie, echo=FALSE, warning=FALSE, message=FALSE,fig.height=5, fig.width=15}
# just for the plots
df_pie1 <- as.data.frame(table(categorized$Category))
colnames(df_pie1) <- c("Category", "count")
# Calculate percentages and label positions
df_pie1 <- df_pie1 %>%
  mutate(
    percentage = count / sum(count) * 100,
    label = paste0(count, "\n(", round(percentage, 1), "%)")
  )

p1 <- ggplot(df_pie1, aes(x = "", y = count, fill = Category)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    color = "black"
  ) +
  scale_fill_manual(values = biocopy_colors) +
  labs(title = paste0("Total peptides: ", nrow(categorized)), fill = "Category") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

p2 <- ggplot(categorized, aes(Total_Mismatches, fill = Category)) +
  geom_bar(position = "dodge") +
  theme_light() +
  scale_fill_manual(values = biocopy_colors) +
  labs(title = "Off-targets Classification by TPT",
       subtitle = "OT - position permissiveness Classification according to the target",
       x = "Edit distance",
       y = "Count") #+
  #theme(legend.position = "bottom")

p1 + p2
```


## Step 3: Estimate _Multi-Features score_

```{r bayesian, include=T, echo=F, comment=NA}

# Bayesian_LR_sensitivity.R: to access LR thresholds
# Bayesian_workflow.R: the main orchestrator -> will be tranformed to this section directly
# utils_exp.R: will test if it is needed at all
# Bayesian_peptide_assessment.R:
# Bayesian_preprocess_experimental_data.R:

source("../MAGEA3_analysis/Multi_features_ranking_Utils.R")
source("../MAGEA3_analysis/Bayesian_peptide_assessment.R")
source("../MAGEA3_analysis/Bayesian_preprocess_experimental_data.R")



```

# Notes

## Experimental evidence
  In the experimental evidence scoring, HLA-ligand Atlas and PEPREP were not used due to the following reasons:

  - _PEPREP_: At the time, data extraction was not possible from peprep and the schema was not stabilitzed enough to track down the information correctly

  - _HLA-ligand atlas_: at the time there was no real solution for the incomplete mapping between allele-tissue-peptide problem



