---
title: "EpiTox: Full annotations"
author: "Hoor Al-Hasani"
date: "2026-03-23"
params:
  epitox: "/Users/hoor.alhasani/Documents/Projects/Quick/MAGEA4/EpiTox_V2.5/MAGEA4/Reports/AnnoTopes_MAGEA4.xlsx" #"~/Documents/Projects/MAGEA3/EpiTox_results/Cutoff_4/Table/HPA_genes_nTPM.xlsx"   
  xscan: "/Users/hoor.alhasani/Documents/Projects/Quick/MAGEA4/EpiTox_V2.5/MAGEA4/Reports/Xscan_MAGEA4.xlsx" #"~/Documents/Projects/MAGEA3/EpiTox_results//Cutoff_4/Table/Xscan_MAGEA3.xlsx"
  iedb_files: "/Users/hoor.alhasani/Documents/Projects/Quick/MAGEA4/EpiTox_V2.5/MAGEA4/PrediTopes/DB_annotation/"  #"~/Documents/Projects/MAGEA3/EpiTox_results/Cutoff_4/PrediTopes/DB_annotation/"
  target_allele: "HLA-A*02:01"
  target: "GVYDGREHTV" #"KVAELVHFL"
  affinity: 51.104 #14.39613
  DIR: "/Users/hoor.alhasani/Documents/Projects/Quick/MAGEA4/EpiTox_V2.5/MAGEA4/Pre_V3/MAGEA4" #"/Users/hoor.alhasani/Documents/Projects/D003/Patent_Paper/paper_materials/02_Data_Analysis/data/MAGEA3" # with prefix
  SCORE_VR6: "/Volumes/lab/03_HighSCORE/P030_05/P030_05_000_Triplicates_results/P030_VD_05_008_PPB444_TPP106​_anabel104/PPB-444_kinetic_table.xlsx"
  SCORE_VR57: "/Volumes/lab/03_HighSCORE/P030_05/P030_05_000_Triplicates_results/P030_VD_05_016_PPB441_TPP96​_anabel104/PPB-441_kinetic_table.xlsx"
  SCORE_VR58: "/Volumes/lab/03_HighSCORE/P030_05/P030_05_000_Triplicates_results/P030_VD_05_015_PPB442_TPP98​_anabel104/PPB-442_kinetic_table.xlsx"
  SaveResults: false
  
output: html_document
---

# Declare libraries

```{r setup, include=FALSE, echo=FALSE, comment="#"}

biocopy_colors = c("#A2C510", "#99CFE9", "#FBB800", "#939597", "#C61E19", "#438D99", "#958BB2", "#6B7B88",
                   "#338232", "#F08000", "#3373A1", "#64686A", "#D14B47", "#98D0BC", "#4F3D7F", "#2C4255")


knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lessR)
library(tidyr)
library(ggpubr)
library(ggplot2)
library(ggrepel)
library(circlize)
library(Peptides)
library(tidyverse)
library(patchwork)
library(ggcorrplot)
library(ComplexHeatmap)


top_annotation <- function(gxp = NA,  to_heatmap = NA, scale = F){

  if(scale){
    gxp_colSums = colMeans(scale(t(to_heatmap)), na.rm = T)
  }else{
    gxp_colSums = colMeans(t(to_heatmap), na.rm = T)
  }
  top_ha = HeatmapAnnotation(
    anchor_status = gxp$anchor_status,
    WT = gxp$Wildtype,
    log2_target_KD_FC = anno_barplot(gxp$log2_target_FC,  gp = gpar( fill = "#6696B9"), height = unit(1.3, "cm")),
    Total = anno_barplot(
      gxp_colSums,
      bar_width = 1, gp = gpar(fill = "#338232", lwd = 0.2, col ="black"),
      height = unit(1.3, "cm")
    ),
    col = list(
      anchor_status = c("anchor_mismatch" = "#F08000", "Intact" = "#438D99",
                        "Both Anchor & Backbone" = "#EAD7A4"),
      WT = c("No" = "#2C4255", "Yes" = "#F0F0F0")
    ),
    gap = unit(c(1,1,2,2,0,0,2), "mm"),
    border = c(Bi_feature_rank =T, Multi_feature_rank =T, Experimental_Evidence =T)
  )
  return (top_ha)
}

bottom_annotation <- function(gxp = NA,  to_heatmap = NA, scale = F){
  top_ha = HeatmapAnnotation(
    Bi_feature_rank = gxp$Bi_feature_rank,
    Multi_feature_rank = gxp$Multi_feature_rank,
    Experimental_Evidence = gxp$Experimental_Evidence,
    col = list(
      Bi_feature_rank = c("High" = "#00508A", "Moderate" = "#B0DAEE", "Low" = "#F4FAFD"),
      Multi_feature_rank = c("High" = "#00508A", "Moderate" = "#B0DAEE", "Low" = "#F4FAFD"),
      Experimental_Evidence = c("High" = "#00508A", "Moderate" = "#B0DAEE", "Very Low" = "#F4FAFD")
    ),
    #gap = unit(c(1,0,2,2,0,0,2), "mm"),
    border = c(Bi_feature_rank =T, Multi_feature_rank =T, Experimental_Evidence =T)
  )
  return (top_ha)
}

```

### Reading input

the input to the model is all the annotations that EpiTox extracted from IEDB. For this target (_MAGEA4_), both HLA-ligand atlas and PEPREP contribution was subset of the IEDB whole annotation. See notes bellow to understand some of the decisions that were made.

The input is:

- the final table from V2.5's `annotopes`: _HPA_genes_nTPM.xlsx_

- the experimental evidence file from the previous step

- _IEDB_ annotated peptides


## Step 1: Estimate _Experimental Evidence_

The experimental evidence is a Baysian-based model to evaluate and summarize the context evidence that is available for each peptide.
the concept is to determine the peptide with its protein ensuring context aware evaluation.


```{r iedb, include=T, echo=F, comment=NA}

# Bayesian_LR_sensitivity.R: to access LR thresholds
# Bayesian_workflow.R: the main orchestrator -> will be tranformed to this section directly
# utils_exp.R: will test if it is needed at all
# Bayesian_peptide_assessment.R:
# Bayesian_preprocess_experimental_data.R:

source("../MAGEA3_analysis/utils_exp.R")
source("../MAGEA3_analysis/Bayesian_peptide_assessment.R")
source("../MAGEA3_analysis/Bayesian_preprocess_experimental_data.R")

files = list.files(path = params$iedb_files,
                   pattern = "*sequence.csv", full.names = T)

length(files) %>% paste(., "files were found!") %>% print()
basename(files) %>% print()

experimental_df = lapply(files, function(f){
  if (file.size(f) == 0) return(NULL)
  first_line <- readLines(f, n = 1, warn = FALSE)
  if (length(first_line) == 0 || all(trimws(first_line) == "")) return(NULL)

  x = read.csv(f) %>%
    select(parent_source_antigen_name, curated_source_antigen.accession, linear_sequence, reference_id,
           qualitative_measure, mhc_class, mhc_allele_name,
           assay_names, disease_names) %>%
    dplyr::rename(
      peptide = linear_sequence,
      study_ref = reference_id,
      disease = disease_names,
      Experimental_method = assay_names,
      Binding = qualitative_measure,
      hla_class = mhc_class,
      hla_allele = mhc_allele_name)
  return(x)

}) %>%
  do.call("rbind",.) %>%
  mutate(uniprot = gsub("\\.\\d+", "", curated_source_antigen.accession),
         disease = gsub("\\[\\'(.+)\\'\\]", "\\1", disease),
         curated_source_antigen.accession = NULL,
         id = paste0(uniprot, "_", peptide),
         uniprot = NULL, peptide = NULL,
         is_normal = case_when(
           disease == "healthy" ~ TRUE,
           TRUE ~ FALSE
         )) %>%
  relocate(id)
```

```{r epitox, include=FALSE, echo=FALSE, comment="#"}
# loading EpiTox's peptide list
# --------------------------------
epitox = openxlsx::read.xlsx(params$epitox) %>%
  select(ensembl_gene_id, uniprot, Gene.Names, peptide, blosum_similarity, mismatch,
         Wildtype, Peptide_HLA_Atlas, affinity, presentation_score, Ranking_score, Rank) %>%
  distinct(.keep_all = T) %>%
  mutate(id = paste0(uniprot, "_", peptide)) %>%
  dplyr::rename()%>%
  distinct(peptide, uniprot, .keep_all = T)

```

### Core analysis

process experimental data, followed by bayesian assessment then saving the results

```{r bayesian, include=FALSE, echo=FALSE, comment="#"}
preprocessed <- preprocess_for_bayesian(epitox, experimental_df,
                                        target_allele = params$target_allele,
                                        target_disease = NULL,
                                        target_class = "I")

# Assess
bayesian_results <- bayesian_peptide_assessment(
  preprocessed,
  target_allele = params$target_allele
)

if (params$SaveResults) generate_assessment_report(bayesian_results, output_prefix = params$DIR)

```

#### Generate plots
```{r bayesian_visualize_exp_pie, echo=FALSE, warning=FALSE, message=FALSE, comment= NA, fig.height=5, fig.width=15}

cols <- c("High" = "#C61E19", "Medium" = "#FDE399",
          "Low" = "#8ECAE6", "Very Low" = "#A2C510")

df_pie1 <- as.data.frame(table(bayesian_results$confidence_level))
colnames(df_pie1) <- c("confidence_level", "count")
p1 <- ggplot(df_pie1, aes(x = "", y = count, fill = confidence_level)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = cols) +
  labs(title = paste0("Total peptides: ", nrow(bayesian_results)), fill = "Confidence level") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

df_pie2 <- as.data.frame(table(bayesian_results$interpretation))
colnames(df_pie2) <- c("interpretation", "count")
p2 <- ggplot(df_pie2, aes(x = "", y = count, fill = interpretation)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = biocopy_colors) +
  labs(title = paste0("Total peptides: ", nrow(bayesian_results)), fill = "Interpretation") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p1 + p2

print(df_pie1)
```


```{r bayesian_visualize_I, echo=FALSE, comment=NA, fig.keep='all', results = 'hide', warning=FALSE, , message=FALSE}
g = ggplot(bayesian_results, aes(affinity/10^5, posterior_prob)) +
  geom_point(size = 1, aes(col = evidence)) +
  theme_light()

#print (g)

g =ggplot(bayesian_results, aes(posterior_prob, affinity*10^-5)) +
  geom_point(size = 1, aes(col = confidence_level)) +
  theme_light() +
  facet_wrap(~confidence_level)

#print(g)
```

## Step 2: TPT

### Target

```{r TPT, include=T, echo=F, comment=NA}
# create TPT -> target pattern/profile template
source("../MAGEA3_analysis/xscan_target_utils.R") # for target functions
source("../MAGEA3_analysis/xscan_OT_utils.R") # for OT functions


#target = data.frame(peptide = params$target, params$affinity)
substitution_matrix <- openxlsx::read.xlsx(params$xscan) %>%
  select(peptide, affinity) %>%
  convert_to_matrix(df = ., wt_peptide = as.character(params$target))

# Run analysis
anchor_results <- analyze_anchor_positions(substitution_matrix, as.numeric(params$affinity), params$target)
anchor_results <- analyze_position_compensation(anchor_results)

```

```{r TPT_visualize_II, echo=FALSE, comment=NA, fig.keep='all', results = 'hide', warning=FALSE, , message=FALSE, fig.height=5, fig.width=15}
L = nchar(params$target)
p1 = ggplot(anchor_results, aes(x = Position, y = Fold_Change_Mean)) +
  geom_col(aes(fill = Category), color = "#2C4255") +  # Changed from Is_Anchor to Category
  geom_hline(yintercept = 1, linetype = "dashed", color = "#A2C510") +
  scale_fill_manual(values = c("Classical Anchor" = "#A2C510",
                               "Permissive" = "#99CFE9",
                               "Flexible Anchor" = "#4F3D7F",
                               "Neutral" = "#F0F0F0"),
                    name = "Position Type") +
  labs(title = "Mean Fold Change in KD by Position",
       x = "Peptide Position",
       y = "Mean Fold Change\n(Mutant KD / WT KD)") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        legend.position = "none") +
  scale_x_continuous(breaks = 1:L)



p2 = ggplot(anchor_results, aes(x = Position, y = Percent_Better_Than_WT)) +
    geom_col(aes(fill = Category), color = "#2C4255") +
    geom_hline(yintercept = 50, linetype = "dashed", color = "#2C4255") +
    scale_fill_manual(values = c("Classical Anchor" = "#A2C510",
                               "Permissive" = "#99CFE9",
                               "Flexible Anchor" = "#4F3D7F",
                               "Neutral" = "#F0F0F0"),
                    name = "Position Type") +
    labs(title = "Position Permissiveness",
       subtitle = "% of substitutions that maintain/improve binding",
       x = "Peptide Position",
       y = "% Better or Equal to WT") +
    theme_light() +
    theme(text = element_text(size = 10),
                          legend.position = "bottom") +
    scale_x_continuous(breaks = 1:L) +
    coord_flip()

p1 +p2

```

### Off-target classification

```{r TPT_OT, include=T, echo=F, comment=NA}
target_classifications = anchor_results %>%
  mutate(WT_AA = strsplit(params$target, "")[[1]]) %>%
  select(Position, WT_AA, Category)

offtargets_data = epitox %>%
  select(uniprot, peptide, Wildtype, blosum_similarity, mismatch,
         affinity, presentation_score) %>%
  distinct(.keep_all = T)

# Perform complete analysis
analysis_results <- analyze_offtargets(params$target, target_classifications, offtargets_data)

# Create comprehensive summary table
summary_table <- create_summary_table(analysis_results) %>%
  mutate(uniprot = gsub(".+\\_", "", id),
         peptide = gsub("\\_.+$", "", id),
         id = NULL) %>%
  relocate(uniprot, peptide) %>%
  arrange(Total_Mismatches)

anchor_categorized <- summary_table %>%
  mutate(
    Category = case_when(
      Anchor_Mismatches > 0 & Exposed_Mismatches > 0 ~ "Both Anchor & Backbone",
      Anchor_Mismatches > 0 & Exposed_Mismatches == 0 ~ "Anchor Only",
      Anchor_Mismatches == 0 & Exposed_Mismatches > 0 ~ "Backbone Only",
      TRUE ~ "Perfect Match"
    )
  )

anchor_categorized %>% head %>% print()
```

```{r TPT_OT_visualize_pie, echo=FALSE, warning=FALSE, message=FALSE,fig.height=5, fig.width=15}
# just for the plots
df_pie1 <- as.data.frame(table(anchor_categorized$Category))
colnames(df_pie1) <- c("Category", "count")
# Calculate percentages and label positions
df_pie1 <- df_pie1 %>%
  mutate(
    percentage = count / sum(count) * 100,
    label = paste0(count, "\n(", round(percentage, 1), "%)")
  )

p1 <- ggplot(df_pie1, aes(x = "", y = count, fill = Category)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    color = "black"
  ) +
  scale_fill_manual(values = biocopy_colors) +
  labs(title = paste0("Total peptides: ", nrow(anchor_categorized)), fill = "Category") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

p2 <- ggplot(anchor_categorized, aes(Total_Mismatches, fill = Category)) +
  geom_bar(position = "dodge") +
  theme_light() +
  scale_fill_manual(values = biocopy_colors) +
  labs(title = "Off-targets Classification by TPT",
       subtitle = "OT - position permissiveness Classification according to the target",
       x = "Edit distance",
       y = "Count") #+
  #theme(legend.position = "bottom")

p1 + p2
```


## Step 3: Estimate _Multi-Features score_

```{r multi_config, include=F, echo=F, comment=NA, warning=FALSE}

source("../MAGEA3_analysis/Multi_features_ranking_Utils.R")
# library(Peptides)
# library(tidyverse)
# #library(corrplot)
# library(ggrepel)

# ============================================================================
# CONFIGURATION
# ============================================================================

# Top N
N_DISAGREEMENTS <- 20
N_VALIDATION <- 30

# Define backbone positions (TCR-facing)
BACKBONE_POSITIONS <- 3:8  # P3-P8 for 9-mers -> this needs to be re-evaluated based on tpt assay
anchor_positions <- sort(anchor_results$Position[anchor_results$Category == "Classical Anchor"])
all_positions <- min(anchor_positions):max(anchor_positions)
BACKBONE_POSITIONS <- setdiff(all_positions, anchor_positions)


# Feature weights (biology-informed)
# Backbone features get higher weight as they face TCR
WEIGHTS <- c(
  gravy = 1.0,
  pI = 1.0,
  charge = 1.0,
  aromaticity = 1.0,
  aliphatic_index = 1.0,
  backbone_gravy = 2.0,      # 2x weight
  backbone_charge = 2.0,      # 2x weight
  backbone_aromaticity = 2.0  # 2x weight
)

# Normalize weights to sum to 1
WEIGHTS <- WEIGHTS / sum(WEIGHTS)

```

```{r}
print(anchor_positions)
print(BACKBONE_POSITIONS)
```

```{r multi_data, include=F, echo=F, comment=NA}
bayesian = bayesian_results %>%
  dplyr::select(peptide_id, confidence_level, evidence_chain) %>%
  dplyr::rename(id = peptide_id)

positions_categories = anchor_categorized %>%
  mutate(anchor_status = case_when(
    Category == "Anchor Only" ~ "anchor_mismatch",
    Category == "Perfect Match" ~ "Intact",
    Category == "Backbone Only" ~ "Intact",
    TRUE ~ Category),
    id = paste0(uniprot, "_", peptide)) %>%
  dplyr::select(id, anchor_status)

epitox = epitox %>%
  dplyr::rename(GeneName = Gene.Names) %>%
  dplyr::select(uniprot, GeneName, peptide, Wildtype, blosum_similarity, mismatch,
         affinity, presentation_score, Ranking_score, Rank) %>%
  mutate(id = paste0(uniprot, "_", peptide),
         #known_peptide = ifelse(peptide %in% known_ot$peptide, "Known", "unkwon"),
         Rank = ifelse(Rank == "Random", "Low", Rank)) %>%
  distinct(id, .keep_all = T) %>%
  dplyr::rename(Bi_features_score = Ranking_score,
                Bi_feature_rank = Rank) %>%
  merge(., positions_categories, by ="id") %>%
  merge(., bayesian, by ="id") %>%
  relocate(anchor_status, .after = "mismatch")

```

```{r multi_features, include=F, echo=F, comment=NA}
#============================================================================
# COMPUTE MULTI-BIOPHYSICAL FEATURES
# ============================================================================

# Target peptide features
target_features <- compute_biophysical_features(params$target)

# All peptides - compute features
# Use group_by(id) with group_modify to handle duplicates properly
peptide_data <- epitox %>%
  group_by(id) %>%
  group_modify(~ {
    features <- compute_biophysical_features(.x$peptide[1])
    bind_cols(.x, features)
  }) %>%
  ungroup()

# Multi-biophysical feature columns
multibiophys_cols <- c("gravy", "pI", "charge", "aromaticity", "aliphatic_index",
                       "backbone_gravy", "backbone_charge", "backbone_aromaticity")

# Calculate scaling parameters from all peptides
scaling_params <- peptide_data %>%
  dplyr::select(all_of(multibiophys_cols)) %>%
  summarise(across(everything(),
                   list(mean = ~mean(., na.rm = TRUE),
                        sd = ~sd(., na.rm = TRUE))))

# Scale peptide features using computed parameters
peptide_data <- peptide_data %>%
  mutate(
    gravy_scaled = apply_scaling(gravy,
                                 scaling_params$gravy_mean,
                                 scaling_params$gravy_sd),
    pI_scaled = apply_scaling(pI,
                              scaling_params$pI_mean,
                              scaling_params$pI_sd),
    charge_scaled = apply_scaling(charge,
                                  scaling_params$charge_mean,
                                  scaling_params$charge_sd),
    aromaticity_scaled = apply_scaling(aromaticity,
                                       scaling_params$aromaticity_mean,
                                       scaling_params$aromaticity_sd),
    aliphatic_index_scaled = apply_scaling(aliphatic_index,
                                           scaling_params$aliphatic_index_mean,
                                           scaling_params$aliphatic_index_sd),
    backbone_gravy_scaled = apply_scaling(backbone_gravy,
                                          scaling_params$backbone_gravy_mean,
                                          scaling_params$backbone_gravy_sd),
    backbone_charge_scaled = apply_scaling(backbone_charge,
                                           scaling_params$backbone_charge_mean,
                                           scaling_params$backbone_charge_sd),
    backbone_aromaticity_scaled = apply_scaling(backbone_aromaticity,
                                                scaling_params$backbone_aromaticity_mean,
                                                scaling_params$backbone_aromaticity_sd)
  )

# Scale target features using the SAME parameters
target_features_scaled <- target_features %>%
  mutate(
    gravy_scaled = apply_scaling(gravy,
                                 scaling_params$gravy_mean,
                                 scaling_params$gravy_sd),
    pI_scaled = apply_scaling(pI,
                              scaling_params$pI_mean,
                              scaling_params$pI_sd),
    charge_scaled = apply_scaling(charge,
                                  scaling_params$charge_mean,
                                  scaling_params$charge_sd),
    aromaticity_scaled = apply_scaling(aromaticity,
                                       scaling_params$aromaticity_mean,
                                       scaling_params$aromaticity_sd),
    aliphatic_index_scaled = apply_scaling(aliphatic_index,
                                           scaling_params$aliphatic_index_mean,
                                           scaling_params$aliphatic_index_sd),
    backbone_gravy_scaled = apply_scaling(backbone_gravy,
                                          scaling_params$backbone_gravy_mean,
                                          scaling_params$backbone_gravy_sd),
    backbone_charge_scaled = apply_scaling(backbone_charge,
                                           scaling_params$backbone_charge_mean,
                                           scaling_params$backbone_charge_sd),
    backbone_aromaticity_scaled = apply_scaling(backbone_aromaticity,
                                                scaling_params$backbone_aromaticity_mean,
                                                scaling_params$backbone_aromaticity_sd)
  )
```


```{r multi_features_normalized_scores, include=T, echo=F, comment=NA}
# ============================================================================
# COMPUTE MULTI-BIOPHYSICAL DISTANCE (ANCHOR STATUS-AWARE)
# ============================================================================

peptide_data <- peptide_data %>%
  rowwise() %>%
  mutate(
    # Get conditional weights for this peptide based on anchor status
    conditional_weights = list(compute_conditional_weights(anchor_status)),

    # Compute multi-biophysical distance using conditional weights
    multibiophys_distance = {
      feature_names <- names(unlist(conditional_weights))
      weights_vec <- unlist(conditional_weights)

      # Calculate weighted distance
      squared_diffs <- sapply(feature_names, function(feat) {
        feat_scaled <- paste0(feat, "_scaled")
        diff <- cur_data()[[feat_scaled]] - target_features_scaled[[feat_scaled]]
        return(diff^2)
      })

      sqrt(sum(weights_vec * squared_diffs))
    }
  ) %>%
  ungroup()

# ============================================================================
# NORMALIZE SCORES TO COMPARABLE 0-1 SCALE
# ============================================================================

multi_features_results <- peptide_data %>%
  mutate(
    # Bi_features is already 0-1 (keep as is)
    #Bi_features_score = Bi_features,

    # Transform multibiophys_distance to 0-1 similarity score
    # (inverted so higher = more similar = higher priority, matching Bi_features direction)
    multibiophys_similarity = 1 - (
      (multibiophys_distance - min(multibiophys_distance)) /
        (max(multibiophys_distance) - min(multibiophys_distance))
    ),

    # Compute disagreement on 0-1 scale
    score_disagreement = abs(Bi_features_score - multibiophys_similarity),

    # Also create ranks for selection/visualization
    rank_Bi_features = rank(-Bi_features_score, ties.method = "min"),
    rank_multibiophys = rank(-multibiophys_similarity, ties.method = "min"),
    rank_disagreement = abs(rank_Bi_features - rank_multibiophys)
  )

# ============================================================================
# CATEGORIZE MULTI-BIOPHYSICAL DISTANCE (k-means, k=3)
# ============================================================================

# Use of the bounded similarity score
set.seed(42)
kmeans_result <- kmeans(multi_features_results$multibiophys_similarity, centers = 3)

multi_features_results <- multi_features_results %>%
  mutate(
    multibiophys_cluster = kmeans_result$cluster,
    # Relabel so HIGH similarity = cluster 1
    multibiophys_category = case_when(
      multibiophys_cluster == which.max(kmeans_result$centers) ~ "High",
      multibiophys_cluster == which.min(kmeans_result$centers) ~ "Low",
      TRUE ~ "Moderate"
    )
  ) %>%
  relocate(multibiophys_similarity, multibiophys_category, .after = "Bi_feature_rank")


if(params$SaveResults) openxlsx::write.xlsx(multi_features_results, paste0(params$DIR, "_PreV3.xlsx"))
```

```{r multi_features_visualize, echo=FALSE, warning=FALSE, message=FALSE,fig.height=5, fig.width=15}
library(ggpubr)

p1 = ggplot(multi_features_results, aes(Bi_features_score, multibiophys_similarity)) +
  geom_point() +
  geom_smooth() +
  stat_cor(method = "spearman") +
  theme_light() +
  ggtitle("Bi-features vs. Multifeatures ranking scores correlation")


p2 = ggplot(multi_features_results) +
  geom_histogram(aes(multibiophys_similarity), fill = biocopy_colors[1], alpha = 0.7, col = "white") +
  geom_histogram(aes(Bi_features_score), fill = biocopy_colors[2], alpha = 0.7, col = "white") +
  labs(x = "Ranking scores distribution", caption = "Blue = Bi features\nGreen = Multibiophys") +
  theme_light() +
  ggtitle("Bi-features vs. Multifeatures ranking scores distribution")

p1 + p2

```

```{r multi_features_visualize_cor, echo=FALSE, warning=FALSE, message=FALSE,fig.height=5, fig.width=15}

cor_matrix <- multi_features_results %>%
  dplyr::select(all_of(multibiophys_cols), affinity, blosum_similarity) %>%
  cor(use = "complete.obs")

# # Visualize correlations
# corrplot(cor_matrix, method = "circle", type = "upper",
#          col = colorRampPalette(c("#99CFE9", "white", "#A2C510"))(200),
#          title = "Multi-Biophysical Feature Correlations",
#          mar = c(0, 0, 2, 0))

p1 = ggcorrplot(cor_matrix,
           method = "circle",
           type = "upper",
           colors = c("#99CFE9", "white", "#A2C510"),
           title = "Multi-Biophysical Feature Correlations",
           ggtheme = theme_minimal())

# Visualize distance distributions by anchor status
p2 = ggplot(multi_features_results, aes(x = anchor_status, y = multibiophys_distance, fill = anchor_status)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.9) +
  labs(
    title = "Multi-Biophysical Distance Distribution by Anchor Status",
    subtitle = "Conditional weighting schemes applied based on mismatch location",
    x = "Anchor Status",
    y = "Multi-Biophysical Distance from Target"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values =c("Both Anchor & Backbone" = "#FBB800",
                              "Intact" = "#99CFE9", "anchor_mismatch" = "#C61E19"))

p1 + p2
```


```{r multi_features_visualize_anchor, echo=FALSE, warning=FALSE, message=FALSE,fig.height=7, fig.width=15}
#head(multi_features_results)

p1 = multi_features_results %>%
  pivot_longer(cols = all_of(multibiophys_cols),
               names_to = "feature",
               values_to = "value") %>%
  ggplot(aes(x = anchor_status, y = value, fill = anchor_status)) +
  geom_boxplot() +
  facet_wrap(~feature, scales = "free_y", ncol = 3) +
  labs(title = "Multi-Biophysical Features by Anchor Status",
       x = "Anchor Status", y = "Feature Value") +
  theme_light() +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("Both Anchor & Backbone" = "#FBB800",
                               "Intact" = "#99CFE9", "anchor_mismatch" = "#C61E19"))




p1 
```

```{r decision_matrix_plot, echo=FALSE, warning=FALSE, message=FALSE,fig.height=5, fig.width=15}
#library(ggpubr)

p1 = ggplot(multi_features_results, aes(Bi_features_score, multibiophys_similarity)) +
  geom_point() +
  geom_smooth() +
  stat_cor(method = "spearman") +
  theme_light() +
  ggtitle("Bi-features, Multifeatures and experimental evidence ranking scores") +
  facet_wrap(~confidence_level, scales = "free")
p1
```


# Get HighScore resutls

<div style="width: 30%;">
| VR | TPP | anabel | 
|----|-----|-----| 
| 4 | 104 | NA |
| 6 | 106 | 104 |
| 57 | 96 | 104 |
| 58 | 98 | 104 |
</div>

```{r vr6, echo=FALSE, warning=FALSE, message=FALSE,fig.height=7, fig.width=15}

score_reference = openxlsx::read.xlsx(params$SCORE_VR6) %>%
  dplyr::rename(peptide = peptide_sequence) %>%
  # aggregate over target
  group_by(peptide) %>%
  summarise(KD = mean(KD, na.rm =T),
            across(everything(), first),  # keep first value of all columns
            .groups = 'drop'
  ) %>%
  ungroup() %>% # not filtering for n because the majority are non-binders, by removing them the number of tested change
  filter (Outcome == "Binder", peptide %in% multi_features_results$peptide, n_slides_QC_passed >1) %>% # first peptide is the QC droped one, the second is the incorrect one (TVAELVQFL)
  select(labguru_id, peptide, KD, n, lev_dist, Outcome, log2_target_FC, n_slides_QC_passed) %>%
  arrange(lev_dist)

```

```{r vr57, echo=FALSE, warning=FALSE, message=FALSE,fig.height=7, fig.width=15}

score_reference_57 = openxlsx::read.xlsx(params$SCORE_VR57) %>%
  dplyr::rename(peptide = peptide_sequence) %>%
  # aggregate over target
  group_by(peptide) %>%
  summarise(KD = mean(KD, na.rm =T),
            across(everything(), first),  # keep first value of all columns
            .groups = 'drop'
  ) %>%
  ungroup() %>% # not filtering for n because the majority are non-binders, by removing them the number of tested change
  filter (Outcome == "Binder", peptide %in% multi_features_results$peptide, n_slides_QC_passed >1) %>% # first peptide is the QC droped one, the second is the incorrect one (TVAELVQFL)
  select(labguru_id, peptide, KD, n, lev_dist, Outcome, log2_target_FC, n_slides_QC_passed) %>%
  arrange(lev_dist)

```

```{r score_annotation, echo=FALSE, warning=FALSE, message=FALSE,fig.height=5, fig.width=15, eval =F}

epitox_annotation = multi_features_results %>%
  merge(., score_reference, by = "peptide") %>%
  relocate(uniprot, peptide) %>%
  relocate(KD, n, lev_dist, Outcome, log2_target_FC, .after = "multibiophys_category")

epitox_annotation_57 = multi_features_results %>%
  merge(., score_reference_57, by = "peptide") %>%
  relocate(uniprot, peptide) %>%
  relocate(KD, n, lev_dist, Outcome, log2_target_FC, .after = "multibiophys_category")


p1 = ggplot(epitox_annotation,
       aes(Bi_features_score, multibiophys_similarity, col = score_disagreement)) +
  geom_point(aes(shape = Wildtype), size = 5) +
  facet_wrap(~confidence_level) +
  theme_light() +
  ggtitle(paste0("Binder OT according to SCORE - VR6 (n = ", nrow(epitox_annotation), ")")) +
  labs(subtitle = "n_slides_QC_passed >1; Outcome = Binder", 
       x = "Bi-feaature scores", y = "Multi-features score",
       caption = "Plot is faceted by experimental evidence\n[0,1] score_disagreement: 0 = agree, 1 = disagree.") +
  theme(legend.position = "bottom")

p2 = ggplot(epitox_annotation_57,
       aes(Bi_features_score, multibiophys_similarity, col = score_disagreement)) +
  geom_point(aes(shape = Wildtype), size = 5) +
  facet_wrap(~confidence_level) +
  theme_light() +
  ggtitle(paste0("Binder OT according to SCORE - VR57 (n = ", nrow(epitox_annotation_57), ")")) +
  labs(subtitle = "n_slides_QC_passed >1; Outcome = Binder", 
       x = "Bi-feaature scores", y = "Multi-features score",
       caption = "Plot is faceted by experimental evidence\n[0,1] score_disagreement: 0 = agree, 1 = disagree.") +
  theme(legend.position = "bottom")


p1 + p2
```
```{r score_prep, echo=FALSE, warning=FALSE, message=FALSE}
# Define all antibodies here — add more entries to scale up
antibody_configs <- list(
  VR6  = list(score_file = params$SCORE_VR6,  label = "VR6"),
  VR57 = list(score_file = params$SCORE_VR57, label = "VR57")#,
  #VR58 = list(score_file = params$SCORE_VR58, label = "VR58")
)

# Load and filter score data for each antibody
score_data <- lapply(antibody_configs, function(cfg) {
  openxlsx::read.xlsx(cfg$score_file) %>%
    dplyr::rename(peptide = peptide_sequence) %>%
    group_by(peptide) %>%
    summarise(KD = mean(KD, na.rm = T),
              across(everything(), first),
              .groups = 'drop') %>%
    filter(Outcome == "Binder",
           peptide %in% multi_features_results$peptide,
           n_slides_QC_passed > 1) %>%
    select(labguru_id, peptide, KD, n, lev_dist, Outcome, log2_target_FC, n_slides_QC_passed) %>%
    arrange(lev_dist) %>%
    mutate(antibody = cfg$label)
})
```


```{r annotation_prep, echo=FALSE, warning=FALSE, message=FALSE, eval=T}
df_epitox_full <- openxlsx::read.xlsx(params$epitox) %>%
  mutate(id = paste0(uniprot, "_", peptide))

COLS <- colnames(df_epitox_full)[c(1:3, 7, 28:ncol(df_epitox_full))]

# Build annotated gxp dataframe for each antibody
gxp_data <- lapply(score_data, function(score_df) {
  annotation <- multi_features_results %>%
    merge(., score_df, by = "peptide") %>%
    relocate(uniprot, peptide) %>%
    relocate(KD, n, lev_dist, Outcome, log2_target_FC, .after = "multibiophys_category")

  df_filtered <- df_epitox_full %>%
    filter(id %in% annotation$id)

  gxp <- annotation %>%
    select(id, Wildtype, mismatch, blosum_similarity, anchor_status, GeneName,
           confidence_level, Bi_feature_rank, multibiophys_category, Outcome, log2_target_FC) %>%
    dplyr::rename(Multi_feature_rank = multibiophys_category,
                  Experimental_Evidence = confidence_level) %>%
    merge(., df_filtered[, COLS], by = "id", all.x = T) %>%
    mutate(Experimental_Evidence = ifelse(Experimental_Evidence == "Medium", "Moderate", Experimental_Evidence),
           genes = sub("\\s.*", "", GeneName)) %>%
    arrange(anchor_status, Wildtype, Bi_feature_rank, Multi_feature_rank, Experimental_Evidence, log2_target_FC) %>%
    distinct(id, .keep_all = T)

  # Separate target row
  target_row <- gxp %>%
    mutate(peptide = gsub(".+\\_", "", id)) %>%
    filter(peptide == params$target) %>%
    mutate(peptide = NULL)

  gxp <- gxp %>%
    mutate(peptide = gsub(".+\\_", "", id)) %>%
    filter(peptide != params$target) %>%
    mutate(peptide = NULL)

  list(gxp = gxp, target = target_row, annotation = annotation)
})

# Shared tissue categories (structure is identical across antibodies)
tissues_prot <- data.frame(tissue = colnames(gxp_data[[1]]$gxp[16:(ncol(gxp_data[[1]]$gxp)-1)])) %>%
  mutate(category = case_when(
    tissue %in% c("adipose.tissue", "skin") ~ "Connective / Integumentary",
    tissue %in% c("adrenal.gland", "thyroid.gland", "parathyroid.gland", "pancreas") ~ "Endocrine",
    tissue %in% c("heart.muscle", "skeletal.muscle", "smooth.muscle") ~ "Muscular",
    tissue %in% c("cerebral.cortex", "choroid.plexus") ~ "Nervous",
    tissue %in% c("lung") ~ "Respiratory",
    tissue %in% c("liver", "gallbladder") ~ "Hepatic / Biliary",
    tissue %in% c("kidney", "urinary.bladder") ~ "Urinary",
    tissue %in% c("spleen", "thymus", "lymph.node", "tonsil", "appendix", "bone.marrow") ~ "Immune",
    tissue %in% c("colon", "duodenum", "esophagus", "rectum", "salivary.gland", "small.intestine", "stomach", "tongue") ~ "Digestive",
    tissue %in% c("breast", "cervix", "endometrium", "epididymis", "fallopian.tube", "ovary", "placenta", "prostate", "seminal.vesicle", "testis") ~ "Reproductive",
    TRUE ~ "Other"
  ))
```

```{r score_annotation_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=15}

annotation_plots <- lapply(names(score_data), function(ab_name) {
  annotation <- multi_features_results %>%
    merge(., score_data[[ab_name]], by = "peptide") %>%
    relocate(uniprot, peptide) %>%
    relocate(KD, n, lev_dist, Outcome, log2_target_FC, .after = "multibiophys_category")

  ggplot(annotation, aes(Bi_features_score, multibiophys_similarity, col = score_disagreement)) +
    geom_point(aes(shape = Wildtype), size = 5) +
    facet_wrap(~confidence_level) +
    theme_light() +
    ggtitle(paste0("Binder OT according to SCORE - ", ab_name, " (n = ", nrow(annotation), ")")) +
    labs(subtitle = "n_slides_QC_passed >1; Outcome = Binder",
         x = "Bi-feature scores", y = "Multi-features score",
         caption = "Plot is faceted by experimental evidence\n[0,1] score_disagreement: 0 = agree, 1 = disagree.") +
    theme(legend.position = "bottom")
})

wrap_plots(annotation_plots)
```

```{r HM, echo=FALSE, warning=FALSE, message=FALSE, fig.height=11, fig.width=15, eval=T}
build_heatmap <- function(ab_name, show_rows = TRUE, show_legend = TRUE) {
  gxp       <- gxp_data[[ab_name]]$gxp
  target_row <- gxp_data[[ab_name]]$target
  expr_mat  <- gxp[, 16:(ncol(gxp)-1)]
  df_fc     <- sweep(log2(expr_mat + 1), 2,
                     log2(as.numeric(target_row[, 16:(ncol(target_row)-1)]) + 1), "-")
  mat       <- as.matrix(df_fc)
  rownames(mat) <- make.unique(paste0(gxp$genes, "_", gsub(".+\\_", "", gxp$id), "_", ab_name))
  col_fun <- colorRamp2(
    c(min(mat, na.rm=T), 0, 1, 2, 4, 6, max(mat, na.rm=T)),
    c("#64686A", "#F6F9E7", "#D6E6D6", "#ADCDAD", "#85B484", "#5C9B5B", "#338232")
  )
  Heatmap(t(mat),
          name = ab_name,
          col = col_fun,
          column_title = ab_name,
          rect_gp = gpar(col = "black", lwd = 0.2),
          row_gap = unit(2, "mm"), border = TRUE,
          row_split = tissues_prot$category,
          show_row_names = show_rows,
          row_names_side = "right",
          row_title_rot = 0,
          show_heatmap_legend = show_legend,
          row_title_gp = gpar(fontsize = 9),
          row_names_gp = gpar(fontsize = 10),
          column_split = gxp$Wildtype,
          cluster_rows = F, cluster_columns = F,
          column_names_gp = gpar(fontsize = 9),
          show_column_dend = F, show_row_dend = F,
          column_names_rot = 45,
          na_col = "gray",
          top_annotation = top_annotation(gxp = gxp, to_heatmap = mat),
          bottom_annotation = bottom_annotation(gxp = gxp, to_heatmap = mat),
          heatmap_legend_param = list(
            direction = "horizontal",
            legend_width = unit(8, "cm"),
            title = paste0("log2 FC - ", ab_name)
          )
  )
}

# For side-by-side display: row names and legend only on last
heatmap_list_display <- lapply(seq_along(names(gxp_data)), function(i) {
  ab_name <- names(gxp_data)[i]
  is_last <- i == length(gxp_data)
  build_heatmap(ab_name, show_rows = is_last, show_legend = is_last)
})

combined_hm <- Reduce("+", heatmap_list_display)
draw(combined_hm,
     column_title = "Confirmed binders by highSCORE",
     merge_legend = TRUE,
     heatmap_legend_side = "bottom",
     annotation_legend_side = "bottom")

if (params$SaveResults) {
  pdf(paste0(params$DIR, "_Binders_GXP_nTPM_tissueSplit_all.pdf"), width = 12, height = 11)
  lapply(names(gxp_data), function(ab_name) {
    hm <- build_heatmap(ab_name, show_rows = TRUE, show_legend = TRUE)
    draw(hm,
         column_title = paste0("Confirmed binders - ", ab_name),
         merge_legend = TRUE,
         heatmap_legend_side = "bottom",
         annotation_legend_side = "bottom")
  })
  dev.off()
  invisible(dev.off())
}
```

```{r binders_table, echo=FALSE, warning=FALSE, message=FALSE}
library(DT)

# Combine all antibody results into one table
binders_combined <- lapply(names(gxp_data), function(ab_name) {
  gxp_data[[ab_name]]$annotation %>%
    mutate(antibody = ab_name)
}) %>%
  bind_rows() %>%
  select(antibody, uniprot, peptide, Wildtype, mismatch, anchor_status,
         blosum_similarity, affinity, Bi_features_score, multibiophys_similarity,
         confidence_level, KD, log2_target_FC, Outcome) %>%
  arrange(antibody, mismatch)

datatable(binders_combined,
          filter = "top",
          rownames = FALSE,
          options = list(pageLength = 5,
                         scrollX = TRUE),
          caption = "Confirmed binders per antibody — filterable by any column")
```

```{r gxp_prep_vr6, echo=FALSE, warning=FALSE, message=FALSE,fig.height=7, fig.width=15, eval=F}
df = openxlsx::read.xlsx(params$epitox) %>%
  mutate(id = paste0(uniprot, "_", peptide)) %>%
  filter(id %in% epitox_annotation$id) %>%
  relocate(id)

COLS = colnames(df)[c(1:3,7,28:ncol(df))]
epitox_gxp = epitox_annotation%>%
  select(id, Wildtype, mismatch, blosum_similarity, anchor_status,
         confidence_level, Bi_feature_rank, multibiophys_category, Outcome, log2_target_FC) %>%
  dplyr::rename(Multi_feature_rank = multibiophys_category,
                Experimental_Evidence = confidence_level) %>%
  merge(., df[,COLS], by = "id", all.x =T) %>%
  mutate(Experimental_Evidence = ifelse(Experimental_Evidence == "Medium", "Moderate", Experimental_Evidence),
         genes = sub("\\s.*", "", Gene.Names)) %>%
  arrange(anchor_status, Wildtype, Bi_feature_rank,
          Multi_feature_rank, Experimental_Evidence, log2_target_FC) %>%
  distinct(id, .keep_all = T) # this was added after the final version since MAGEA2 had two arrows in the gxp -> the binders are 36 not 37

tissues_prot = data.frame(tissue = colnames(epitox_gxp[16:(ncol(epitox_gxp)-1)])) %>%
  mutate(category = case_when(
    tissue %in% c("adipose.tissue", "skin") ~ "Connective / Integumentary",
    tissue %in% c("adrenal.gland", "thyroid.gland", "parathyroid.gland", "pancreas") ~ "Endocrine",
    tissue %in% c("heart.muscle", "skeletal.muscle", "smooth.muscle") ~ "Muscular",
    tissue %in% c("cerebral.cortex", "choroid.plexus") ~ "Nervous",
    tissue %in% c("lung") ~ "Respiratory",
    tissue %in% c("liver", "gallbladder") ~ "Hepatic / Biliary",
    tissue %in% c("kidney", "urinary.bladder") ~ "Urinary",
    tissue %in% c("spleen", "thymus", "lymph.node", "tonsil","appendix", "bone.marrow") ~ "Immune", #"Immune / Lymphatic",
    tissue %in% c("colon", "duodenum", "esophagus", "rectum", "salivary.gland", "small.intestine", "stomach", "tongue") ~ "Digestive",
    tissue %in% c("breast", "cervix", "endometrium", "epididymis", "fallopian.tube", "ovary", "placenta", "prostate", "seminal.vesicle", "testis") ~ "Reproductive",
    #tissue %in% c("appendix", "bone.marrow") ~ "Immune / Hematopoietic",
    TRUE ~ "Other"
  ))

target = epitox_gxp %>%
  mutate(peptide = gsub(".+\\_", "", id)) %>%
  filter (peptide == params$target) %>%
  mutate(peptide = NULL)


epitox_gxp = epitox_gxp %>%
  mutate(peptide = gsub(".+\\_", "", id)) %>%
  filter(peptide != params$target) %>%
  mutate(peptide = NULL)
```

```{r gxp_prep_vr57, echo=FALSE, warning=FALSE, message=FALSE,fig.height=7, fig.width=15, eval=F}
df = openxlsx::read.xlsx(params$epitox) %>%
  mutate(id = paste0(uniprot, "_", peptide)) %>%
  filter(id %in% epitox_annotation_57$id) %>%
  relocate(id)

COLS = colnames(df)[c(1:3,7,28:ncol(df))]
epitox_gxp_57 = epitox_annotation_57 %>%
  select(id, Wildtype, mismatch, blosum_similarity, anchor_status,
         confidence_level, Bi_feature_rank, multibiophys_category, Outcome, log2_target_FC) %>%
  dplyr::rename(Multi_feature_rank = multibiophys_category,
                Experimental_Evidence = confidence_level) %>%
  merge(., df[,COLS], by = "id", all.x =T) %>%
  mutate(Experimental_Evidence = ifelse(Experimental_Evidence == "Medium", "Moderate", Experimental_Evidence),
         genes = sub("\\s.*", "", Gene.Names)) %>%
  arrange(anchor_status, Wildtype, Bi_feature_rank,
          Multi_feature_rank, Experimental_Evidence, log2_target_FC) %>%
  distinct(id, .keep_all = T) 

tissues_prot_57 = data.frame(tissue = colnames(epitox_gxp_57[16:(ncol(epitox_gxp_57)-1)])) %>%
  mutate(category = case_when(
    tissue %in% c("adipose.tissue", "skin") ~ "Connective / Integumentary",
    tissue %in% c("adrenal.gland", "thyroid.gland", "parathyroid.gland", "pancreas") ~ "Endocrine",
    tissue %in% c("heart.muscle", "skeletal.muscle", "smooth.muscle") ~ "Muscular",
    tissue %in% c("cerebral.cortex", "choroid.plexus") ~ "Nervous",
    tissue %in% c("lung") ~ "Respiratory",
    tissue %in% c("liver", "gallbladder") ~ "Hepatic / Biliary",
    tissue %in% c("kidney", "urinary.bladder") ~ "Urinary",
    tissue %in% c("spleen", "thymus", "lymph.node", "tonsil","appendix", "bone.marrow") ~ "Immune", #"Immune / Lymphatic",
    tissue %in% c("colon", "duodenum", "esophagus", "rectum", "salivary.gland", "small.intestine", "stomach", "tongue") ~ "Digestive",
    tissue %in% c("breast", "cervix", "endometrium", "epididymis", "fallopian.tube", "ovary", "placenta", "prostate", "seminal.vesicle", "testis") ~ "Reproductive",
    #tissue %in% c("appendix", "bone.marrow") ~ "Immune / Hematopoietic",
    TRUE ~ "Other"
  ))

target_57 = epitox_gxp_57 %>%
  mutate(peptide = gsub(".+\\_", "", id)) %>%
  filter (peptide == params$target) %>%
  mutate(peptide = NULL)


epitox_gxp_57 = epitox_gxp_57 %>%
  mutate(peptide = gsub(".+\\_", "", id)) %>%
  filter(peptide != params$target) %>%
  mutate(peptide = NULL)
```

```{r hm_vr6, echo=FALSE, warning=FALSE, message=FALSE,fig.height=11, fig.width=15, eval=F}
expr_mat = epitox_gxp[, 16:(ncol(epitox_gxp)-1)]

df <- sweep(log2(expr_mat+1), 2, log2(as.numeric(target[, 16:(ncol(target)-1)])+1), "-") # took the isoform with highest gxp
mat = as.matrix(df)

#cat("min:", min(mat, na.rm = TRUE), "\n")
#cat("max:", max(mat, na.rm = TRUE), "\n")
#cat("class:", class(mat), "\n")
#cat("all NA?", all(is.na(mat)), "\n")

rownames(mat) = paste0(epitox_gxp$genes, "_", gsub(".+\\_", "", epitox_gxp$id))
peptides = gsub(".+\\_", "", epitox_gxp$id)
# Define your custom breakpoints for better granularity above zero
col_prot_fun = colorRamp2(
  c(min(mat, na.rm = TRUE), 0, 1, 2, 4, 6, max(mat, na.rm = TRUE)),
  c("#64686A", "#F6F9E7", "#D6E6D6", "#ADCDAD", "#85B484", "#5C9B5B", "#338232")
)
prot_top_ha = top_annotation(gxp = epitox_gxp, to_heatmap = mat)
prot_bottom_ha = bottom_annotation(gxp = epitox_gxp, to_heatmap = mat)

hpa_ch_plot = mat %>%
  t() %>%
  Heatmap(name = "Gene expression (nTPM)",
          col = col_prot_fun,
          rect_gp = gpar(col = "black", lwd = 0.2),
          row_gap = unit(2, "mm"), border = TRUE,
          row_split = tissues_prot$category,
          row_title_rot = 0,
          row_title_gp = gpar(fontsize = 9),
          row_names_gp = gpar(fontsize = 10),
          column_split = epitox_gxp$Wildtype,
          
          cluster_rows = F,
          cluster_columns = F,
          column_names_gp = gpar(fontsize = 9),
          show_column_dend = F,
          show_row_dend = F,
          column_names_rot = 45,
          
          na_col = "gray",
          top_annotation = prot_top_ha,
          bottom_annotation = prot_bottom_ha,
          heatmap_legend_param = list(
            direction = "horizontal",
            legend_width = unit(15, "cm"),
            at = c(min(mat, na.rm = TRUE), 0, 1, 2, 4, 6, 8, max(mat, na.rm = TRUE)),
            labels = c(round(min(mat, na.rm = TRUE)), "0", "1", "2", "4", "6", "8",
                       round(max(mat, na.rm = TRUE))),
            title = "log2 FC"
          ),
  )

# generate heatmap
draw(hpa_ch_plot, column_title = "Confirmed binders according to highSCORE - VR6",
           merge_legend = TRUE, heatmap_legend_side = "bottom",
           annotation_legend_side = "bottom")

if(params$SaveResults){
  pdf(paste0(params$DIR,"/Binders_GXP_nTPM_tissueSplit_VR6.pdf"), width = 14, height = 12)
  draw(hpa_ch_plot, column_title = "Confirmed binders according to highSCORE",
     merge_legend = TRUE, heatmap_legend_side = "bottom",
     annotation_legend_side = "bottom")
dev.off()}
```
```{r hm_vr57, echo=FALSE, warning=FALSE, message=FALSE,fig.height=11, fig.width=15, eval=F}
expr_mat_57 = epitox_gxp_57[, 16:(ncol(epitox_gxp_57)-1)]

df <- sweep(log2(expr_mat_57+1), 2, log2(as.numeric(target_57[, 16:(ncol(target_57)-1)])+1), "-") 
mat_57 = as.matrix(df)

rownames(mat_57) = paste0(epitox_gxp_57$genes, "_", gsub(".+\\_", "", epitox_gxp_57$id))
peptides_57 = gsub(".+\\_", "", epitox_gxp_57$id)
# Define your custom breakpoints for better granularity above zero
col_prot_fun = colorRamp2(
  c(min(mat_57, na.rm = TRUE), 0, 1, 2, 4, 6, max(mat_57, na.rm = TRUE)),
  c("#64686A", "#F6F9E7", "#D6E6D6", "#ADCDAD", "#85B484", "#5C9B5B", "#338232")
)
prot_top_ha_57 = top_annotation(gxp = epitox_gxp_57, to_heatmap = mat_57)
prot_bottom_ha_57 = bottom_annotation(gxp = epitox_gxp_57, to_heatmap = mat_57)

hpa_ch_plot_57 = mat_57 %>%
  t() %>%
  Heatmap(name = "Gene expression (nTPM)",
          col = col_prot_fun,
          rect_gp = gpar(col = "black", lwd = 0.2),
          row_gap = unit(2, "mm"), border = TRUE,
          row_split = tissues_prot_57$category,
          row_title_rot = 0,
          row_title_gp = gpar(fontsize = 9),
          row_names_gp = gpar(fontsize = 10),
          column_split = epitox_gxp_57$Wildtype,
          
          cluster_rows = F,
          cluster_columns = F,
          column_names_gp = gpar(fontsize = 9),
          show_column_dend = F,
          show_row_dend = F,
          column_names_rot = 45,
          
          na_col = "gray",
          top_annotation = prot_top_ha_57,
          bottom_annotation = prot_bottom_ha_57,
          heatmap_legend_param = list(
            direction = "horizontal",
            legend_width = unit(15, "cm"),
            at = c(min(mat_57, na.rm = TRUE), 0, 1, 2, 4, 6, 8, max(mat_57, na.rm = TRUE)),
            labels = c(round(min(mat_57, na.rm = TRUE)), "0", "1", "2", "4", "6", "8",
                       round(max(mat_57, na.rm = TRUE))),
            title = "log2 FC"
          ),
  )

# generate heatmap
draw(hpa_ch_plot_57, column_title = "Confirmed binders according to highSCORE - VR57",
           merge_legend = TRUE, heatmap_legend_side = "bottom",
           annotation_legend_side = "bottom")

if(params$SaveResults){
  pdf(paste0(params$DIR,"_Binders_GXP_nTPM_tissueSplit_VR57.pdf"), width = 14, height = 12)
  draw(hpa_ch_plot, column_title = "Confirmed binders according to highSCORE",
     merge_legend = TRUE, heatmap_legend_side = "bottom",
     annotation_legend_side = "bottom")
dev.off()}
```
# Notes

## Experimental evidence
  In the experimental evidence scoring, HLA-ligand Atlas and PEPREP were not used due to the following reasons:

  - _PEPREP_: At the time, data extraction was not possible from peprep and the schema was not stabilitzed enough to track down the information correctly

  - _HLA-ligand atlas_: at the time there was no real solution for the incomplete mapping between allele-tissue-peptide problem



